---
title: "GE Aerospace Case Study"
author: "Samantha Erne, Jake Holroyd, Jersie Thomas"
editor: visual
format:
  revealjs: 
    theme: [default]
    logo: https://www.ge.com/news/themes/custom/genewsroom/logo.svg
    footer: "ISA 401- Miami University"
    fontcolor: "white"
    backgroundcolor: "navy"
---

## Problem Statement

The problem that GE Aerospace is facing is that aircrafts using GE gas turbine engines are experiencing unscheduled outages

-   The engines are nearing the end of their remaining useful life faster than expected

## Case Study Outline {background-image="https://www.ge.com/content/dam/gepower-new/global/en_US/images/gas-new-site/products/gas-turbines/7f/hero-7f-gas-turbine.png" background-size="500px" background-position="50% 75%"}

For this case study, our goal was to create a model that would accurately predict the remaining useful life (RUL) of GE Engines

-   This will allow for more accurate scheduling and maintenance

## Our Process

In order to solve the RUL problem, we took the following steps:

-   Merge and Clean the Data

-   Create Statistical Summaries

-   Visualize the Data

-   Create the Model

-   Host an App

## Step 1: Merge and Clean the Data

We started the process by merging the data from the 7 CSV files into one R data frame. Then we...

-   Checked that the data was tidy, technically correct, and consistent
    -   Tweaked column names and imputed null fields to achieve a clean data set
    
    -   Corrected typos
    
-   Exported the cleaned data into a CSV file

## Step 1: Merge and Clean the Data
Below is a snapshot of the cleaned dataset, pulled directly from the CSV

```{r}
library(tidyverse)
Tech_Correct = read_csv("/Users/sammieerne/Downloads/ge_cleaned_data_corrected.csv") 
head(Tech_Correct)
```
## Step 1: Merge and Clean the Data

Below is a summary of the cleaned and merged dataset

```{r}
summary(Tech_Correct)
```

## Step 1: Merge and Clean the Data

Below is a graphic showing that we were able to impute all null values in the data

```{r}
pacman::p_load(DataExplorer)
plot_missing(Tech_Correct)
```

## Step 2: Create Statistical Summaries

Next, we created a Jupyter Notebook and utilized autoviz tools to create statistical summaries of the data

-   These summaries helped us to determine how variables were correlated
    -   We used these insights to create our final predictive model
  
## Step 2: Create Statistical Summaries {background-image="https://github.com/sammieerne/project_code/blob/main/project_1.PNG?raw=true" background-size="600px" background-position="50% 65%"}

This visualization is a heat map of correlation between the variables

## Step 2: Create Statistical Summaries {background-image="https://github.com/sammieerne/project_code/blob/main/project_2.PNG?raw=true" background-size="900px" background-position="50% 65%"}

This visualization shows which variables are constants

## Step 2: Create Statistical Summaries {background-image="https://github.com/sammieerne/project_code/blob/main/project_3.PNG?raw=true" background-size="900px" background-position="50% 65%"}

This visualization provides a summary of the data set

## Step 2: Create Statistical Summaries {background-image="https://github.com/sammieerne/project_code/blob/main/project_4.PNG?raw=true" background-size="600px" background-position="50% 70%"}

This visualization shows the interaction between RUL and flight cycle

## Step 2: Create Statistical Summaries {background-image="https://github.com/sammieerne/project_code/blob/main/project_5.PNG?raw=true" background-size="600px" background-position="50% 70%"}

This visualization shows the interaction between RUL and distance


## Step 3: Visualize the Data

Next, we used Tableau to create dashboards showcasing our clean and merged data

-   These dashboards helped  show relationships between predictors in the data set

-   The dashboards also provided insights on GE's clients

## Step 3: Visualize the Data {background-image="https://github.com/sammieerne/project_code/blob/main/tableau_1.png?raw=true" background-size="500px" background-position="50% 80%"}

The visualizations below show distance and flight cycles plotted against RUL, lines of best fit, and average values by customer


## Step 3: Visualize the Data {background-image="https://github.com/sammieerne/project_code/blob/main/tableau_2.png?raw=true" background-size="500px" background-position="50% 80%"}

This set of visualizations shows correlation scatter plots with lines of best fit

## Step 3: Visualize the Data {background-image="https://github.com/sammieerne/project_code/blob/main/tableau_3.png?raw=true" background-size="500px" background-position="50% 80%"}

This set of visualizations shows animations of where the engines are travelling and how distance is affecting RUL

## Step 4: Create the Model

Then, we worked to create a model that would accurately predict RUL's of GE engines

-   We decided how to aggregate each field in the data so that we could have one record per engine

-   We ran regressions based on what we thought the most predictive variables were

    -   We tested models with different variables until we decided on our best model
    
## Step 4: Create the Model {background-image="https://github.com/sammieerne/project_code/blob/main/validation_model.png?raw=true" background-size="450px" background-position="50% 80%"}

-   Our final model had a mean R-square of 0.6220, and RMSE of 30, and 8 predictors

## Step 5: Host an App

Finally, we used Hugging Spaces to host a permanent app to showcase our predictive model



## Our Insights {background-image="https://github.com/sammieerne/project_code/blob/main/model_validation1.png?raw=true" background-size="900px" background-position="50% 75%"}

We were able to utilize the CSV files provided to create a model with less variables and more accurate predictions of RUL. 



## Closing Remarks {background-image="https://mms.businesswire.com/media/20220718005465/en/1515995/5/GE_Aerospace_Logo.jpg" background-size="700px" background-position="50% 75%"}

Special Thanks to:

- GE Aerospace

- Dr. Fadel Megahed

